---
title: "Perfecto-Sevilla_Quiz2.rmd"
author: "Covi Perfecto, Luis Carlo Sevilla"
date: "7/29/2021"
output: html_document
---
#### Question 1: 
An article in the Journal of Sound and Vibration ["Measurement of Noise-Evoked Blood Pressure by Means of Averaging Method: Relation between Blood Pressure Rise and PSL" (1991, Vol. 151(3), pp. 383-394)] described a study investigating the relationship between noise exposure and hypertension. The following data are representative of those reported in the article.

```{r, echo= FALSE}
bpr=c(1,0,1,2,5,1,4,6,2,3,5,4,6,8,4,5,7,9,7,6) 
psl=c(60,63,65,70,70,70,80,90,80,80,85,89,90,90,90,90,94,100,100,100)
BPR_PSL= data.frame(psl,bpr)
BPR_PSL
```

***A.Draw a scatter diagram of y (blood pressure rise in millimeters of mercury) versus x (sound pressure level in decibels). Does a simple linear regression model seem reasonable in this situation?***



It is reasonable to use the Linear Regression Model in analyzing this problem because the result that the research wants to answer is the relationship between the two variables--- the sound pressure levels(dB) and blood pressure rise in millimeter of mercury. 



We can utilize the library of ***ggplot2*** to create an accurate regression model of the sound pressure levels(dB) and blood pressure rise in millimeter of mercury. We can do so by calling the library of ***ggplot2*** and appointing the x and y values and using the **ggplot(data=MyDataFile, aes(x,y))** function. 
```{r, echo=TRUE}
library(ggplot2)
scatterBPR_PSL=ggplot(data=BPR_PSL, aes(bpr,psl)) +   
  geom_point()
  
scatterBPR_PSL = scatterBPR_PSL+ geom_point(size=2) +
  xlab("PSL (dB)") +  
  ylab("Blood Pressure Rise (mmHg)") +   
  ggtitle("Measurement of Noise-Evoked Blood Pressure") 
scatterBPR_PSL
```

***B. Fit the simple linear regression model using least squares. Find an estimate of σ2.***



**Step 1:**



To fit a simple linear regression model, we can use the equation of the fitted or estimated regression line:

$$
\hat{y} = \hat{\beta}_0 + \hat{\beta}_1x
$$

where $\hat{\beta}_0$ and $\hat{\beta}_1x$ are represented (according to the least squares) as:

$$
\hat{\beta}_0 = \bar{y} - \hat{\beta}_1\bar{x}
$$

$$
\hat{\beta}_1 = \frac{S_{xx}}{S_{xy}}
$$ 

where $S_{xx} = \sum_{i = 1}^{n} x^2_i - \frac{(\sum_{i = 1}^{n} x_i)^2}{n}$, $S_{xy} = \sum_{i = 1}^{n} x_iy_i - \frac{(\sum_{i = 1}^{n} x_i)(\sum_{i = 1}^{n} y_i)}{n}$,  $\bar{y} = (1/n) \sum_{i = 1}^{n} y_i$, and $\bar{x} = (1/n) \sum_{i = 1}^{n} x_i$ 



**Step 2:**



Let us get the required values to solve for the linear regression line using the functions of R:



The sample size:



$n = 20$ ; 




$$\sum_{i = 1}^{20} x_i = 1656$$

```{r, echo = TRUE}
sumX = sum(psl); sumX
```
$$\sum_{i = 1}^{20} y_i = 86$$
```{r, echo = TRUE}
sumY = sum(bpr); sumY
```
$$\bar{x} = 82.8$$
```{r, echo = TRUE}
barX = mean(psl); barX
```
$$\bar{y} = 4.3$$
```{r, echo = TRUE}
barY = mean(bpr); barY
```
$$\sum_{i = 1}^{20} x_i^2 = 140176$$
```{r, echo = TRUE}
sumXsqr = sum(psl^2); sumXsqr
```
$$\sum_{i = 1}^{20} y_i^2 = 494$$
```{r, echo = TRUE}
sumYsqr =sum(bpr^2); sumYsqr
```

$$\sum_{i = 1}^{20} x_1 y_i = 7654$$
```{r, echo = TRUE}
sumXY = sum(bpr*psl); sumXY
```

$$S_{xx} = \sum_{i = 1}^{20} x^2_i - \frac{(\sum_{i = 1}^{20} x_i)^2}{n} = 140176 - \frac{1656^2}{20}= 3059.2$$
```{r, echo = TRUE}
Sxx = sumXsqr- ((sumX^2)/20); Sxx
```


$$S_{xy} = \sum_{i = 1}^{n} x_iy_i - \frac{(\sum_{i = 1}^{n} x_i)(\sum_{i = 1}^{n} y_i)}{n} = 7654 - \frac{(1656)(86)}{20} = 533.2 $$

```{r, echo = TRUE}
Sxy = sumXY - ((sumX*sumY)/20); Sxy
```

**Step 3:**



Therefore, the least squares estimates of the slope and intercepts are:


$$
\hat{\beta}_1 = \frac{S_{xy}}{S_{xx}} = \frac{533.2}{3059.2} = 0.1742939
$$ 
```{r, echo = TRUE}
betaHat1 = Sxy/Sxx; betaHat1
```

and

$$
\hat{\beta}_0 = \bar{y} - \hat{\beta}_1\bar{x} = 4.3 - (0.1742939)82.6 = -10.13154 
$$

```{r, echo = TRUE}
betHat0 = barY - (betaHat1*barX); betHat0
```

**Step 4:**



The fitted simple linear regression model (with the coefficients reported to three decimal places) is
$$
\hat{y} = -10.132 + 0.174x
$$
We can also check if our coefficients are correct by utilizing the ***coef*** function of ggplot:
```{r}
coeftest = lm(bpr~psl)
coef(coeftest)
```

Here is the scatter plot with the fitted linear regression

```{r, echo=FALSE}
library(ggplot2)
scatterBPR_PSL=ggplot(data=BPR_PSL, aes(psl,bpr)) +   
  geom_point()
  
scatterBPR_PSL = scatterBPR_PSL+ geom_point(size=2) +
  xlab("PSL (dB)") +  
  ylab("Blood Pressure Rise (mmHg)") +   
  ggtitle("Measurement of Noise-Evoked Blood Pressure") +
  geom_smooth(method = "lm")
scatterBPR_PSL
```




**Step 5:**
Let us now find the estimate of $\sigma^2$:




To find the estimate of $\sigma^2$, we need to look for the value of $SS_E$  where 


$$SS_E = \sum_{i = 1}^{n} (y_i - \hat{y})^2$$
We can also easily check the Squared Sum due to Errors by using the ***dataBase$resduals*** function of R:
```{r, echo = TRUE}
SSe = sum(scatterBPR_PSL$residuals^2); SSe
```

$$SS_E = 0
$$

finally, we can solve for $\sigma^2$ by

$$
\sigma^2 = \frac{SS_E}{n-2}
$$

Since $SS_E = 0$ that means that we can estimate $\sigma^2$ will also be equal to $0$

**Interpretation:** 

If $SS_E = 0$---which we can also assume that $\sigma^2 = 0$--- means that all of the observations are in the regression line and there is no residual in data given.




***C. Find the predicted mean rise in blood pressure level associated with a sound pressure level of 85 decibels.***



Since we already established the equation of the linear regression we can easily get the predicted mean rise in blood pressure level (y) associated with a sound pressure level of 85 decibels.

$$
\hat{y} = -10.132 + 0.174x = -10.132 + 0.174(85) 
$$
Therefore,
$$
\hat{y} = 4.658
$$


